Measuring security performance in the Android ecosystem

### 2015-06-04 ccs.tex
Update related work

\section{Related work}
The update process for apps, security fixes and OS upgrades also needs to be secure.
Unfortunately, package management systems designed to provide secure updates have been found to contain vulnerabilities~\cite{Cappos2008} and many software update systems fail to authenticate the connection between the device and the update server or do not authenticate the downloaded binaries~\cite{Bellissimo2006}.
Android does authenticate update binaries and Google Play downloads them over a secure connection~\cite{Viennot2014}.
In this paper we have analysed four critical vulnerabilities in the Android app update mechanism: APK unsigned shorts, APK unchecked name, APK duplicate file and Fake ID.
Other work has demonstrate complex and subtle errors exist in the Android app update process.
For example, the process can be exploited to allow apps to gain privilege through `Pileup' vulnerabilities by registering for new permissions before the update which creates that permission is installed~\cite{Xing2014}.

\interfootnotelinepenalty=10000 %To stop the lifehacker footnote below from spanning multiple pages and breaking
User-Agent strings have been used to investigate the timeliness of web browser updates, with at most 80\% of Firefox users running the most recent version~\cite{Frei2008}.
The same analysis was used to show that Chrome's use of silent updates seems to increase uptake of upgrades~\cite{Duebendorfer2010} with 97\% of users running the latest version within 3 weeks of release.
By way of comparison, Android's update process is manual.
The user is notified an update exists, but further action is required, including downloading the update and rebooting the phone to enable installation.
The phone must have sufficient charge to perform the update and the device itself is rendered inoperable during the update process, two factors which might prevent or delay the update process from taking place.
In our data we are unable to determine why a device is not updated. 
It is possible that many updates arrive at handsets, but are simply not installed.
Anecdotal evidence at least suggests that it is the lack of updates rather the lack of installation which is the major problem at present. Further work is required to tease these numbers apart.
Partly this is the result of the fact that an operating system update is being installed and so a reboot is required, but Chrome installs the new version side by side with the old one and switches the next time it is restarted.
The same technique would be more difficult on phones with limited storage space (as many cheap Android phones have barely enough space to install just the update) but is a plausible improvement for more high-end devices.
Google is deploying the same silent update technique through Google Play Services\footnote{\href{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}} which automatically installs updates for core Google components of Android, this also bypasses the device manufacturer and network operator.

Security in depth is also a useful strategy.
In this regard, iOS provides additional safeguards beyond those used in Android, including a pre-distribution review, mandatory code-signing by the manufacturer, and (with the important exception of ROP-based attacks~\cite{Wang2013a}) the technical prohibition of dynamic code loading by an app.
These features, combined with mandatory access controls, has resulted in a lower level of malware affecting iOS when compared to Android~\cite{Felt2011}.

@@

Rather than fixing critical vulnerabilities, security can be obtained by detecting malicious apps and preventing their installation or execution.
Detection strategies include Risk\-Ranker, which classified 3\,281 out of 118\,318 apps (2.8\%) as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown malware, an infection rate of 0.6\% across multiple markets~\cite{Grace2012a}.
Droid\-Ranger also analysed apps finding 148 out of 182\,823 apps (0.08\%) to be malicious across multiple markets of which 29 were previously unknown~\cite{Zhou2012a}.
%DroidRanger: It used permission-based behavioural fingerprinting which looked at the permissions of known malware and heuristic-based filtering -- dynamic loading of both Dalvik and native code.
A common technique used by attackers is to include malicious code in repackaged popular apps. 
An\-Darwin uses this insight to detect similar apps, and found 169 out of 265\,359 of all apps studied (0.06\%) were malicious clones~\cite{Crussell2013}.
%AnDarwin: It used clustering based on semantic vectors derived from the program dependence graphs to detect similar apps.

@@

\todo{Work these references into the text}
\cite{Nappa2015}
\cite{Zhang2014}
\cite{Lindorfer2014}
\cite{Arp2014}

>>>>

\section{Related work}
We assume updates make security better but, the update process for apps, security fixes and OS upgrades also needs to be secure.
Unfortunately, package management systems designed to provide secure updates have been found to contain vulnerabilities~\cite{Cappos2008} and many software update systems fail to authenticate the connection between the device and the update server or do not authenticate the downloaded binaries~\cite{Bellissimo2006}.
Android does authenticate update binaries and Google Play downloads them over a secure connection~\cite{Viennot2014}.
In this paper we have analysed four critical vulnerabilities in the Android app update mechanism: APK unsigned shorts, APK unchecked name, APK duplicate file and Fake ID.
Other work has demonstrated complex and subtle errors exist in the Android app update process.
For example, the process can be exploited to allow apps to gain privilege through `Pileup' vulnerabilities by registering for new permissions before the update which creates that permission is installed~\cite{Xing2014}.

We used \da\ to record data on the OS versions of Android installed on devices, Nappa et.\ al. performed a similar analysis using WINE to record data on installed versions of Windows client programs and used cleaned NVD data rather than \avo\ for vulnerability data, they did not assign client applications scores~\cite{Nappa2015}.

\interfootnotelinepenalty=10000 %To stop the lifehacker footnote below from spanning multiple pages and breaking
We use User-Agent strings from Rwandan internet traffic to examine the version distribution of Android OS versions.
User-Agent strings have been used to investigate the timeliness of web browser updates, with at most 80\% of Firefox users running the most recent version~\cite{Frei2008}.
The same analysis was used to show that Chrome's use of silent updates seems to increase uptake of upgrades~\cite{Duebendorfer2010} with 97\% of users running the latest version within 3 weeks of release.
By way of comparison, Android's update process is manual.
The user is notified an update exists, but further action is required, including downloading the update and rebooting the phone to enable installation.
The phone must have sufficient charge to perform the update and the device itself is rendered inoperable during the update process, two factors which might prevent or delay the update process from taking place.
In our data we are unable to determine why a device is not updated. 
It is possible that many updates arrive at handsets, but are simply not installed.
Anecdotal evidence at least suggests that it is the lack of updates rather the lack of installation which is the major problem at present and \S\ref{sec:security_scoring:results} supports this. Further work is required to tease these numbers apart.
Partly this is the result of the fact that an operating system update is being installed and so a reboot is required, but Chrome installs the new version side by side with the old one and switches the next time it is restarted.
The same technique would be more difficult on phones with limited storage space (as many cheap Android phones have barely enough space to install just the update) but is a plausible improvement for more high-end devices.
Google is deploying the same silent update technique through Google Play Services\footnote{\href{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}} which automatically installs updates for core Google components of Android, this also bypasses the device manufacturer and network operator.

As well as supplying security updates promptly, the impact of vulnerabilities can be reduced through security in depth.
In this regard, iOS provides additional safeguards beyond those used in Android, including a pre-distribution review, mandatory code-signing by the manufacturer, and (with the important exception of ROP-based attacks~\cite{Wang2013a}) the technical prohibition of dynamic code loading by an app.
These features, combined with mandatory access controls, has resulted in a lower level of malware affecting iOS when compared to Android~\cite{Felt2011}.

@@

Rather than fixing critical vulnerabilities, security can be obtained by detecting malicious apps and preventing their installation or execution.
Detection strategies include Risk\-Ranker, which classified 3\,281 out of 118\,318 apps (2.8\%) as risky of which 718 (22\%) were malware and 322 (10\%) were previously unknown malware, an infection rate of 0.6\% across multiple markets~\cite{Grace2012a}.
Droid\-Ranger also analysed apps finding 148 out of 182\,823 apps (0.08\%) to be malicious across multiple markets of which 29 were previously unknown~\cite{Zhou2012a}.
A common technique used by attackers is to include malicious code in repackaged popular apps. 
An\-Darwin uses this insight to detect similar apps, and found 169 out of 265\,359 of all apps studied (0.06\%) were malicious clones~\cite{Crussell2013}.
Other approaches have made further use of malware dependency graphs~\cite{Lindorfer2014} or tried to extract further semantic information to avoid malware being able to avoid detection~\cite{Zhang2014} but simpler approaches based on extracting strings from the binaries can be more effective~\cite{Arp2014} in terms of false positives and false negatives.
There is a lack of ground truth data, a danger in training algorithms based on APK files which other algorithms have already found which might lead to new malware being missed and DroidChameleon showed that existing AntiVirus apps could be fooled by simple permutations of malicious APKs~\cite{Rastogi2013}.

@@
